### 2019-08-20 big-data/paper/paper.tex
Add thesis and big data seminar skeleton

\documentclass[12pt,a4paper]{scrartcl}
\usepackage[protrusion=true,expansion=true]{microtype}
\linespread{1.25}

\usepackage{float}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage[bookmarks=true,hidelinks=true,breaklinks=true]{hyperref}
\usepackage{mathptmx}
\usepackage{cleveref}
\usepackage{enumitem}
\usepackage[backend=biber,natbib=true]{biblatex}
\usepackage[scaled=.92]{helvet}
\bibliography{references.bib}

\newtheorem{lemma1}{Lemma}

\begin{document}
\title{}
\author{Arne Beer, MN 6489196, University of Hamburg}
\date{20.08.2019}

\maketitle

\section{Introduction}

Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test Test test

\subsection{Conventional Hashing}

\section{Simhash}

\subsection{Weights and Hamming Distance}

\section{Bit order}

\subsection{Ordering}

\subsection{Volatility Heap}

%\begin{figure}[H]
%    \includegraphics[scale=0.33]{./gfx/vosh-performance.png}
%    \centering
%    \caption{Performance results comparing random comparison search and volatility order based search.}\label{fig:vosh-performance}
%\end{figure}

\subsection{Probabilistic Simhash Matching}

\section{Conclusion}

\printbibliography
\end{document}