### 2019-08-20 thesis/thesis/chapter/003_acronyms.tex

\chapter*{Acronyms}
\begin{acronym}
    \acro{api}[API]{Application Programming Interface}
    \acro{cet}[CET]{Central European Time}
    \acro{cest}[CEST]{Central European Summer Time}
    \acro{cpu}[CPU]{Central Processing Unit}
    \acro{cvcs}[CVCS]{Centralized Version Control Systems}
    \acro{dbscan}[DBSCAN]{Density-Based Spatial Clustering of Applications with Noise}
    \acro{dst}[DST]{daylight saving time}
    \acroplural{dst}[DSTs]{daylight saving times}
    \acro{eu}[EU]{European Union}
    \acro{fs}[FS]{file system}
    \acro{gdpr}[GDPR]{General Data Protection Regulation}
    \acro{gb}[GB]{Gigabyte}
    \acroplural{gb}[GBs]{Gigabytes}
    \acro{http}[HTTP]{Hypertext Transfer Protocol}
    \acro{iana}[IANA]{Internet Assigned Numbers Authority}
    \acro{io}[I/O]{Input/Output}
    \acro{json}[JSON]{JavaScript Object Notification}
    \acro{mit}[MIT]{Massachusetts Institute of Technology}
    \acro{orm}[ORM]{Object-Relational Mapping}
    \acro{os}[OS]{Operating System}
    \acroplural{os}[OSs]{Operating Systems}
    \acro{sha1}[SHA-1]{Secure Hash Algorithm 1}
    \acro{ssh}[SSH]{Secure Shell}
    \acro{sql}[SQL]{Structured Query Language}
    \acro{tb}[TB]{Terrabytes}
    \acro{url}[URL]{Uniform Resource Locator}
    \acroplural{url}[URLs]{Uniform Resource Locators}
    \acro{utc}[UTC]{Coordinated Universal Time}
    \acro{vcs}[VCS]{version control system}
    \acroplural{vcs}[VCSs]{version control systems}
\end{acronym}