### 2019-08-20 thesis/thesis/chapter/evaluation/employee-punchcard-table.tex

\begin{landscape}
    \begin{table}[]
        \centering
        \captionof{table}{Employee cluster evaluation.}\label{tbl:employee-cluster-evaluation}
        \begin{tabular}{lllll}
            \toprule
            Nickname      & Employer     & Main projects                & Work related & Over 90\% work related \\
            \midrule
                          &              &                              &              &                        \\
            brettfo       & Microsoft    & visualfsharp                 & yes          & no                     \\
            aputinski     & Salesforce   & Salesforce                   & yes          & yes                    \\
            alxhub        & Google       & Angular                      & yes          & yes                    \\
            MatrixFrog    & Google       & Google Projects              & yes          & yes                    \\
            ryanemerson   & Red Hat      & Infinispan                   & yes          & no                     \\
            garagatyi     & Red Hat      & Eclipse and related projects & yes          & yes                    \\
            eternoendless & PrestaShop   & PrestaShop                   & yes          & yes                    \\
            initvector    & vanilla      & vanilla                      & yes          & yes                    \\
            kyhavlov      & HashiCorp    & HashiCorp                    & yes          & no                     \\
            XenoPhex      & CloudFoundry & CloudFoundry                 & yes          & yes                    \\
            gjoranv       & unknown      & vespa-engine/vespa           & probably     & yes                    \\
            doolse        & Equella      & Equella                      & yes          & yes                    \\
            leplatrem     & Mozilla      & Mozilla/Kinto                & yes          & yes                    \\
            StrongMonkey  & Rancher Labs & Rancher Labs                 & yes          & no                     \\
            lukaseder     & jOOQ         & jOOQ                         & yes          & yes                    \\
            DaazKu        & vanilla      & vanilla                      & yes          & yes                    \\
            brettcannon   & Microsoft    & Microsoft/Python             & yes          & yes                    \\
            isidorn       & Microsoft    & Microsoft Projects           & yes          & yes                    \\
            jackhorton    & Microsoft    & Microsoft Projects           & yes          & yes                    \\
            glasserc      & Mozilla      & Mozilla/Kinto                & yes          & yes                    \\
            nickwei84     & CloudFoundry & CloudFoundry                 & yes          & yes                    \\
            \bottomrule
        \end{tabular}
    \end{table}
\end{landscape}