### 2019-08-20 thesis/thesis/chapter/evaluation/leasure-punchcard-table.tex

\begin{landscape}
    \begin{table}[]
        \centering
        \captionof{table}{Leisure time developer cluster evaluation.}\label{tbl:leasure-time-cluster-evaluation}
        \begin{tabular}{llllll}
            \toprule
            Nickname      & Employer      & Main projects exists               & Work related    & Over 90\% work related & Commits rarely \\
            \midrule
                          &               &                                    &                 &                        &                \\
            craigbarratt  & unknown       & backuppc                           & unknown         & unknown                & yes            \\
            hackebrot     & Mozilla       & Own repos, open-source             & no              & no                     & no             \\
            itamarnet     & unknown       & Open-source                        & no              & no                     & yes            \\
            JamesNK       & unknown       & Own repos                          & no              & no                     & no             \\
            kpreid        & unknown       & Own repos                          & no              & no                     & yes            \\
            vrasneur      & unknown       & Own repos                          & no              & no                     & yes            \\
            chrissimpkins & SourceFoundry & Own repos, SourceFoundry           & yes             & no                     & no             \\
            klauscfhq     & Cookfood HQ   & Own repos                          & no              & no                     & no             \\
            knowthelist   & unknown       & Own repos                          & no              & no                     & yes            \\
            larskanis     & Comcard       & Open-source                        & no              & no                     & no             \\
            NickeManarin  & unknown       & Own repos                          & no              & no                     & yes            \\
            obilodeau     & GoSecure      & GoSecure                           & yes             & no                     & no             \\
            piotrmurach   & unknown       & Own repos                          & no              & no                     & no             \\
            hauke         & unknown       & Open-source                        & unknown         & unknown                & yes            \\
            rspeele       & unknown       & Own repos                          & no              & no                     & yes            \\
            DylanC        & PhotoFlare    & PhotoFlare                         & yes             & yes                    & yes            \\
            Gitmaninc     & unknown       & SecWiki                            & no              & no                     & yes            \\
            lgreski       & unknown       & Own repos                          & no              & no                     & yes            \\
            vitaut        & Facebook      & Open-source                        & no              & no                     & no             \\
            wikimatze     & MyHammer AG   & Own repos, open-source             & no              & no                     & yes            \\
            schaal        & unknown       & Own repos                          & no              & no                     & yes            \\
            \bottomrule
        \end{tabular}
    \end{table}
\end{landscape}