### 2019-08-20 thesis/thesis/thesis.tex

\documentclass{thesis}

\usepackage{lipsum}
\usepackage[utf8]{inputenc}
\usepackage[ngerman,english]{babel}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{adjustbox}
\usepackage{pdflscape}
\usepackage[toc,page]{appendix}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{lmodern}
\usepackage{float}
\usepackage{sidecap}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{tabularcalc}
\usepackage{todonotes}
\usepackage{hyperref}
\usepackage{minted}
\usepackage{siunitx}
\usepackage{acronym}
\usepackage{subfig}
\usepackage{tabularx}
\usepackage{setspace}
\usepackage[customcolors]{hf-tikz}
\usepackage{url}
\usepackage{csquotes}
\usepackage{booktabs}
\usepackage[T1]{fontenc}
\usepackage[alldates=long]{biblatex}
\addbibresource{thesis.bib}
\graphicspath{{images/}}
\pgfplotsset{compat=1.12}

\title{Privacy implications of exposing Git meta data}
\author{Arne Beer}

\pagenumbering{roman}
\begin{document}

\input{./chapter/000_title.tex}
\cleardoublepage{}

\input{./chapter/001_abstract.tex}
\clearpage

\input{./chapter/002_quote.tex}

% Set table of contents depth
\setcounter{secnumdepth}{1}
\setcounter{tocdepth}{1}
{\small \tableofcontents}

\input{./chapter/003_acronyms.tex}

\chapter{Introduction}
\pagenumbering{arabic}
\input{./chapter/introduction/introduction.tex}

\input{./chapter/004_related_works.tex}

\chapter{Data Source and Aggregation}\label{data}
This chapter will attend to the collection of required data as stated in Section~\ref{attack-goals}.
At first the \ac{vcs} \emph{Git} will be presented and its functionalities explained.
The actual source of the data \emph{Github} will then be evaluated in terms of the amount of ground truth and availability.
At last the methodology used for aggregation and exploration of Github will be explained.

\input{./chapter/data/git.tex}

\input{./chapter/data/github.tex}

\input{./chapter/data/aggregator.tex}

\chapter{Attack Implementations}~\label{implementation}
After collecting all necessary data as shown in Chapter~\ref{data}, this chapter now attends to the analysis of this data.
The approach for several attacks, as listed in Section~\ref{attack-goals} will be presented and their attack goals recapitulated.
The possible applications for the gained knowledge will be stated and the implementation of and requirements to the respective algorithm will be explained for each attack.

\input{./chapter/implementation/miss-out.tex}

\input{./chapter/implementation/punchcard.tex}

\clearpage
\input{./chapter/implementation/travel.tex}

\chapter{Evaluation and Interpretation}\label{evaluation}
In the previous chapter, the implementation of several possible attacks which could be performed on the gathered data has been shown.
This chapter will now attend to the evaluation of all results gained from these attacks.
I will present the extracted information from each algorithm and compare it to the real-world ground truth.
This information will be then be explained and audited in terms of precision and reliability.

\input{./chapter/evaluation/miss-out.tex}

\input{./chapter/evaluation/punchcard.tex}

\input{./chapter/evaluation/travel.tex}

\input{./chapter/005_conclusion.tex}

\begin{appendices}
\chapter{Leisure time cluster evaluation}
\input{./chapter/evaluation/leasure-punchcard-table.tex}
\end{appendices}

\begingroup
    \footnotesize
    \listoffigures
    \let\clearpage\relax
    \listoflistings{}
    \listoftables
\endgroup

\renewcommand*{\bibfont}{\footnotesize}
\printbibliography{}

\input{./chapter/999_statutory_declaration.tex}

\end{document}